Models of VOR learning in MHC knockout mice

### 2013-02-10 Notes/VORmodel.tex
started OG multistate section

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[height=2cm]{cascade.svg}}\label{fig:cascade_model}\hspace{0.5cm}
  \item\aligntop{\includegraphics[height=1.7cm]{multistate.svg}}\label{fig:multistate_model}\hspace{0.5cm}
  \item\aligntop{\includegraphics[height=1.7cm]{binary.svg}}\label{fig:binary_model}
 \end{myenuma}
 \end{center}
  \caption{Transition probabilities for different models.
  (\ref{fig:cascade_model}) In the cascade model, the transition probabilities decay geometrically with a parameter $x$ (see \cite{Fusi2005cascade}).
  (\ref{fig:multistate_model}) In the multistate model the transition probabilities for potentiation/depression are all equal and it is parameterised by these two values.
  (\ref{fig:binary_model}) The two-state model is parameterised by the two transition probailities.} \label{fig:models}
\end{figure}

@@

\resultsfig{cascade_short}{Simulation results for cascade model with short pre-training ($rt_\text{pre}=50$)}

\resultsfig{cascade_long}{Simulation results for cascade model with long pre-training ($rt_\text{pre}=150$)}

@@

\resultsfig{multistate_weak}{Simulation results for multistate model with weak training ($\Delta f=-0.1$)}

\resultsfig{multistate_strong}{Simulation results for multistate model with strong training ($\Delta f=-0.4$)}

@@

@@

@@

\resultsfig{pooled_plenty}{Simulation results for pooled resource model with light resource depletion ($q\lmin/q\lmax=0.75$)}

\resultsfig{pooled_scarce}{Simulation results for pooled resource model with heavy resource depletion ($q\lmin/q\lmax=0.25$)}

>>>>

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[height=2cm]{cascade.svg}}\label{fig:cascade_model}\hspace{0.5cm}
  \item\aligntop{\includegraphics[height=1.7cm]{multistate.svg}}\label{fig:multistate_model}\hspace{0.5cm}
  \item\aligntop{\includegraphics[height=1.7cm]{binary.svg}}\label{fig:binary_model}
 \end{myenuma}
 \end{center}
  \caption{Transition probabilities for different models.
  (\ref{fig:cascade_model}) In the cascade model, the transition probabilities decay geometrically with a parameter $x$ (see \cite{Fusi2005cascade}).
  (\ref{fig:multistate_model}) In the multistate model the transition probabilities for potentiation/depression are all equal and it is parameterised by these two values.
  (\ref{fig:binary_model}) The two-state model is parameterised by the two transition probabilities.} \label{fig:models}
\end{figure}

@@

\resultsfig{cascade_short}{Simulation results for the cascade model with short pre-training ($rt_\text{pre}=50$)}

\resultsfig{cascade_long}{Simulation results for the cascade model with long pre-training ($rt_\text{pre}=150$)}

@@

\resultsfig{multistate_weak}{Simulation results for the multistate model with weak training ($\Delta f=-0.1$)}

\resultsfig{multistate_strong}{Simulation results for the multistate model with strong training ($\Delta f=-0.4$)}

@@

\subsubsection{OG multistate model}\label{sec:multistate_lin}

In this section, we will consider the multistate model as originally defined in \cite{amit1994learning}, \ie with linearly varying synaptic weight:
%
\begin{equation}\label{eq:multistateLinWeight}
  \w_i = \frac{2i-M-1}{M-1}.
\end{equation}
%
The equilibrium distribution, \eqref{eq:mutltieq}, still applies.
However, now the rate of change of our learning metric, \eqref{eq:learning}, will be proportional to the sum of the net fluxes between adjacent states:
%
\begin{equation}\label{eq:multiLinFlux}
  \begin{aligned}
    \Phi &= \sum_{i=1}^{M-1} \eq_{i+1} f'{}\dep q\dep - \eq_i f'{}\pot q\pot 
         &= \sum_{i=1}^{M-1} \eq_i (\alpha-\alpha') f'{}\dep q\dep \\
         &= \frac{1-\alpha^{M-1}}{1-\alpha^M} \, (\alpha-\alpha') f'{}\dep q\dep.
  \end{aligned}
\end{equation}
%

First, consider the wild-type, for which $q\pot=q\dep=q$.
Without pre-training:
%
\begin{equation}\label{eq:multiLinWTnopre}
  \Phi = -2\Delta fq\,\frac{M-1}{M},
\end{equation}
%
where it's worth remembering that $\Delta f<0$.
With pre-training:
%
\begin{equation}\label{eq:multiLinWTpre}
\begin{aligned}
  \Phi &= -4\Delta f q \, \frac{(1+2\Delta f)^{M-1} - (1-2\Delta f)^{M-1}}
          {(1+2\Delta f)^M - (1-2\Delta f)^M} \\
       &= -{4\Delta f q}\,\frac{M-1}{M} + \CO(\Delta f)^2.
\end{aligned}
\end{equation}
%

Now, consider the mutant, for which $q\pot=\beta q\dep=q$, $\beta<1$.
Without pre-training:
%
\begin{equation}\label{eq:multiLinKOnopre}
  \Phi = -2\Delta f q\,\frac{1-\beta^{M-1}}{1-\beta^M}.
\end{equation}
%
With pre-training:
%
\begin{equation}\label{eq:multiLinKOpre}
\begin{aligned}
  \Phi &= -4\Delta f q \, \frac{(1+2\Delta f)^{M-1} - \beta^{M-1}(1-2\Delta f)^{M-1}}
          {(1+2\Delta f)^M - \beta^M(1-2\Delta f)^M} \\
       &= -4\Delta f q\,\frac{1-\beta^{M-1}}{1-\beta^M} + \CO(\Delta f)^2.
\end{aligned}
\end{equation}
%

@@

We made a number of assumptions in our analysis so far.
Before we declare this model to be falsified, we should think about what would happen if we relaxed these assumptions and explored a larger parameter space.

First, we assumed that the wild type had symmetric potentiation and depression.
Relaxing this assumption would correspond modelling the wild-type like the mutant, but with a different value for $\beta$.
We will still have $\beta\ko<\beta\wt$, and looking at \eqref{eq:binKOnopre} tell us that we would still have faster learning in the mutant.

Secondly, we assumed that there would be equal rates of potentiation and depression in the untrained situation.
However, relaxing this assumption merely changes \eqref{eq:binKOnopre} to
%
\begin{equation}\label{eq:binKOnoprediff}
  \Phi = -\frac{\Delta f q}{f\dep+\beta f\pot}.
\end{equation}
%
This will not change our conclusions either.

As we didn't need the pre-training paradigm to rule out this model, we don't need to look at removing the symmetry between gain-increase and decrease.

@@

\resultsfig{pooled_plenty}{Simulation results for the pooled resource model with light resource depletion ($q\lmin/q\lmax=0.75$)}

\resultsfig{pooled_scarce}{Simulation results for the pooled resource model with heavy resource depletion ($q\lmin/q\lmax=0.25$)}
